Vigilantism and Institutions:\\ Understanding Attitudes toward Lynching in Brazil

### 2022-10-28 article/article.tex
updated article

% English
\usepackage[english]{babel}
\usepackage[english]{isodate}
% \cleanlookdateon

@@

% Bibliography
\usepackage{natbib}
\bibliographystyle{chicago-ff}
\makeatletter
% Remove comma after author
\setcitestyle{aysep={}}
\patchcmd{\NAT@citex}
	  {\@citea\NAT@hyper@{%
		 \NAT@nmfmt{\NAT@nm}%
		 \hyper@natlinkbreak{\NAT@aysep\NAT@spacechar}{\@citeb\@extra@b@citeb}%
		 \NAT@date}}
	  {\@citea\NAT@nmfmt{\NAT@nm}%
	   \NAT@aysep\NAT@spacechar\NAT@hyper@{\NAT@date}}{}{}
	\patchcmd{\NAT@citex}
	  {\@citea\NAT@hyper@{%
		 \NAT@nmfmt{\NAT@nm}%
		 \hyper@natlinkbreak{\NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi}%
		   {\@citeb\@extra@b@citeb}%
		 \NAT@date}}
	  {\@citea\NAT@nmfmt{\NAT@nm}%
	   \NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi\NAT@hyper@{\NAT@date}}
	  {}{}
% Patch case where name and year are separated by aysep
\patchcmd{\NAT@citex}
  {\@citea\NAT@hyper@{%
     \NAT@nmfmt{\NAT@nm}%
     \hyper@natlinkbreak{\NAT@aysep\NAT@spacechar}{\@citeb\@extra@b@citeb}%
     \NAT@date}}
  {\@citea\NAT@nmfmt{\NAT@nm}%
   \NAT@aysep\NAT@spacechar\NAT@hyper@{\NAT@date}}{}{}
% Patch case where name and year are separated by opening bracket
\patchcmd{\NAT@citex}
  {\@citea\NAT@hyper@{%
     \NAT@nmfmt{\NAT@nm}%
     \hyper@natlinkbreak{\NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi}%
       {\@citeb\@extra@b@citeb}%
     \NAT@date}}
  {\@citea\NAT@nmfmt{\NAT@nm}%
   \NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi\NAT@hyper@{\NAT@date}}
  {}{}
\makeatother

@@

% Abstract
\begin{abstract}
\doublespacing \noindent Why do people support extrajudicial violence? In two survey experiments with respondents in Brazil, we examine which characteristics of lynching scenarios garner greater support for lynching and whether providing different types of information about lynching reduces support for it. We find that people often do support community members to take vengeance. In particular, our analysis finds that people strongly support the use of extrajudicial violence by families of victims against men who sexually assault and murder women and children. We also find that criminal punishment and the threat of vendettas reduce support, but appeals to the human rights of victims have zero effect on support for lynchings. 
\vspace{.25cm}

\noindent \textbf{Keywords:} extralegal violence, vigilantism, lynching, Brazil, crime 
\vspace{.25cm}

@@

The regulation of violence is a crucial function of the state. When states
control violence effectively, markets, politics, and civil society can flourish
\citep{besley2011pillars, north2009violence}. However, states still struggle to
maintain the monopoly of force in much of the Global South. Scholars have
written extensively about how rebel organizations build support to challenge
government institutions, yet we know little about what motivates local
extrajudicial violence in the developing world \citep{bateson2020politics,
smith2000new}. This omission is significant not only because local violence is
pervasive in weak states, but also because it can be even more costly than war-
or terrorism-related violence \citep{blair2017predicting}.

One of the most serious forms of extrajudicial violence is lynching, which can
be defined as ``incidents of physical violence committed by large numbers of
private citizens against one or more individuals accused of having committed a
``criminal'' offense, whether or not this violence resulted in the death of the
victim(s)'' \citep[645]{godoy2004justice}. Although lynchings occur in more
than one hundred countries in all regions of the world
\citep[33]{jung2020lynching}, Latin America has been particularly affected by a
sharp increase in vigilante violence. In recent years, countries like Brazil
and Venezuela have experience nearly one execution every two days, and lynching
episodes have also been reported in Guatemala, Mexico, Peru, and other
countries in the region \citep{cruz2019determinants, godoy2004justice,
barbara2015vigilantes}. These alarming figures suggest that, despite being
dismissed by policymakers as irrational and senseless, vigilantism enjoys
considerable approval in Latin America \citep{berg2011globalizing,
goldstein2005flexible}. This raises a number of puzzles: under which conditions
do citizens perceive lynchings as a legitimate response to crime? Are some
groups seen as more ``deserving'' of lynchings that others? And how can we
reduce pro-vigilantism attitudes in young, unequal democracies?

In this study, we run two survey experiments and use primary qualitative data
to understand attitudes toward lynchings in Brazil. First, we employ a conjoint
experiment with eight attributes to evaluate in which circumstances citizens
perceive lynchings as a justifiable punishment for crime. We also included
open-ended questions for respondents to elaborate on their decision process.
Finally, we use an information provision experiment to test the efficacy of
three strategies to reduce individual support for lynchings.

Our empirical approach offers several distinct advantages. First, the data we
collected come from the same level of analysis as the phenomenon itself: local
residents. Second, online surveys are less susceptible to social desirability
bias, and survey experiments help to elicit sensitive information---an
especially important benefit in researching this topic \citep{grimm2010social}.
Third, the experimental nature of the approach allows us to precisely estimate
the size of effects and compare different theoretical explanations
simultaneously. Finally, we can also control for confounding, which is a
potential concern in previous research on vigilantism.

@@

@@

We report marginal means in our main analysis instead of average marginal
component effects (AMCEs). \citet{leeper2018subgroup} show that AMCEs can be
misleading in subgroup comparisons as model results are sensitive to the choice
of reference categories in interactions. In contrast, marginal means allow for
a clear description of quantities of interest, in our case preferences about
lynching, while also allowing for easy comparisons between groups of
respondents. In a forced-choice experiment such as this one, a marginal-means
estimate of 0.5 indicates that respondents are indifferent to this attribute
vis-à-vis other attributes. When the coefficient is lower than 0.5, respondents
dislike profiles with this attribute. Conversely, when the point estimate is
higher than 0.5, respondents prefer profiles containing a given attribute.

@@

Respondents selected male offenders (0.548, SE = 0.003, $p$-value $<$ 0.001) as
preferred lynching victims more often than female offenders (0.438, SE = 0.004,
$p$-value $<$ 0.001). We found identical results in our subgroup analyses, in
which we disaggregated our sample according to respondents' gender, income,
race, support for the death penalty, and views on the police and judiciary. In
all cases, men and teenage boys were more likely to be selected as more
justified victims of lynching. Compared with the attributes we discuss below,
the gender of the crime perpetrator had one of the largest effects in our
experiment. With regard to the age of the crime perpetrator, respondents
tolerated the lynching of adults (0.520, SE = 0.005, $p$-value $<$ 0.001) more
than attacks against teenagers (0.498, SE = 0.005, $p$-value = 0.650) or the
elderly (0.483, SE = 0.005, $p$-value $<$ 0.001). More precisely, subjects were
indifferent toward teenagers but reacted positively to lynching of adults and
negatively to lynching of elderly people.

When we analyzed the effect of the race of the crime perpetrator, in contrast
with the American experience, respondents favored the lynching of white
perpetrators (0.512, SE = 0.006, $p$-value = 0.050) the most. Of all the racial
identities that we included in the experiment, Black people were the least
likely to be chosen as a more justified lynching victim (0.490, SE = 0.006,
$p$-value = 0.092), although the 95 percent confidence intervals overlap with
the remaining categories. The effect is also robust for most subgroups,
including white respondents. Asian respondents, however, rated fellow Asians as
their least preferred choice. Taken together, the results provide experimental
evidence that support for lynchings in Brazil does not resemble the typical
racial patterns scholars have observed in the United States
\citep{dray2003hands,seguin2019national,obert2018keeping}. These findings are
consistent with recent research on vigilantism in Haiti
\citep{jung2020lynching} and with journalistic observations. They further
suggest that the US experience with lynchings might be distinct from
vigilantism in other places and times \citep{oliveira2016mob}.

The residency of the perpetrator is not a significant driver of support for
lynching. Subjects were indifferent as to whether the crime perpetrator lives
in the area (0.498, SE = 0.004, $p$-value = 0.668) or in another neighborhood
(0.502, SE = 0.004, $p$-value = 0.668). In this sense, it is possible that
subjects see lynchings not as a means to protect the area in which they live,
which people commonly suppose vigilante groups do, but rather as retribution in
specific cases.

The next attribute is type of offense. Rape was by far the most significant
attribute in our model (0.719, SE = 0.009, $p$-value $<$ 0.001), followed by
murder (0.608, SE = 0.006, $p$-value $<$ 0.001) and molestation (0.538, SE =
0.006, $p$-value $<$ 0.001). It is not surprising that murder and rape were the
most important offenses to respondents. Overall, we found that respondents do
not believe that property crimes, such as car theft (0.351, SE = 0.009,
$p$-value $<$ 0.001) and pickpocketing (0.314, SE = 0.006, $p$-value $<$
0.001), provide justification for lynching. When we analyzed the data by
groups, we also saw that poorer respondents, whose household income is less
than R\$3,000 (US\$650) per month, are more likely to support the lynching of
individuals who commit a property crime.

For age, we observed that respondents were particularly concerned about
children as crime victims (0.571, SE = 0.007, $p$-value $<$ 0.001), as this
attribute is strongly correlated with support for lynchings. Respondents also
thought lynchings are more justified when the victim is a teenager (0.505, SE =
0.007, $p$-value = 0.474), an elderly person (0.479, SE = 0.006, $p$-value $<$
0.001), or an adult (0.464, SE = 0.006, $p$-value $<$ 0.001).

When we analyzed the gender of the crime victim, we found that respondents
support lynchings more strongly when the victim is a woman or teenage girl
(0.511, SE = 0.004, $p$-value = 0.002) and less strongly when the lynching
victim is a man or teenage boy (0.489, SE = 0.004, $p$-value = 0.002). The
difference here is smaller than when we considered the gender of the crime
perpetrator, but as in the previous estimation the 95 percent confidence
intervals of each attribute did not overlap.

Lastly, and related to this, we examined who respondents believed are more
legitimate perpetrators of lynchings. Respondents considered the family to be
the most justified in carrying out a lynching (0.534, SE = 0.007, $p$-value $<$
0.001). Again, the intervals here did not overlap with those of any other
category. Bystanders (0.505, SE = 0.007, $p$-value = 0.450), gangs (0.502, SE =
0.007, $p$-value = 0.815), and neighbors (0.492, SE = 0.007, $p$-value = 0.262)
do not increase or decrease respondents attitudes about lynching. If the police
carry out a lynching, that reduced the perceived justification (0.467, SE =
0.007, $p$-value $<$ 0.001). In sum, respondents did not believe that lynchings
should be carried out by the state but did believe they should be used as a
tool for individual or family retribution.

@@

\begin{table}[ht] \centering
  \caption{Average treatment effects for experiment 3}
  \label{tab:exp03main}
\begin{tabular}{@{\extracolsep{3pt}}lD{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3} }
\\[-1.8ex]\hline \\[-1.8ex]
\\[-1.8ex] & \multicolumn{4}{c}{\textbf{Lynching Support}\vspace{.5cm}} \\
\\[-1.8ex] & \multicolumn{1}{c}{(1)} & \multicolumn{1}{c}{(2)} & \multicolumn{1}{c}{(3)} & \multicolumn{1}{c}{(4)}\\
\hline \\[-1.8ex]
 Constitution and penal code & -4.509^{**} &  &  &  \\
  & (1.805) &  &  &  \\
  Human rights &  & -1.571 &  &  \\
  &  & (1.801) &  &  \\
  Vendettas &  &  & -3.156^{*} &  \\
  &  &  & (1.879) &  \\
  Combined treatments &  &  &  & -3.023^{**} \\
  &  &  &  & (1.493) \\
  Constant & 40.823^{***} & 40.823^{***} & 40.823^{***} & 40.823^{***} \\
  & (1.293) & (1.293) & (1.293) & (1.293) \\
 N & \multicolumn{1}{c}{1,114} & \multicolumn{1}{c}{1,173} & \multicolumn{1}{c}{1,092} & \multicolumn{1}{c}{2,215} \\
\hline \\[-1.8ex]
\multicolumn{5}{l}{$^{*}$p $<$ .1; $^{**}$p $<$ .05; $^{***}$p $<$ .01} \\
\multicolumn{5}{l}{Robust standard errors in parentheses.} \\
\end{tabular}
\end{table}
\normalsize

Table \ref{tab:exp03main} summarizes our main results for this experiment. We find that when respondents are reminded of legal punishments associated with lynchings, their support for lynching decreases by about 11 percent. This result is consistent with the idea that respondents will support an activity less if they perceive it to be more costly. The next treatment group was told the human rights message. The result is not statistically significant and indicates that appeals to individual liberties and the rule of law are likely to fall short. This finding has practical implications, as many campaigns against violent crimes are conducted by human rights groups, which tend to emphasize this type of message. However, participants do not generally believe that offenders do, in fact, have human rights. Accordingly, we have little reason to believe that such messaging is effective. By contrast, Treatment 1 reminds respondents that perpetrators of lynching potentially face a high personal cost from doing so.

Our last treatment, which raises the possibility of sparking vendettas, has a large, negative effect on support for lynchings. Mentioning that lynchings trigger cycles of violence has a negative effect of about 8 percent when compared with the control group. While we expected that the coefficient for this treatment condition would be larger than that for the treatment on legal punishments, the means of both groups are not statistically different from each other ($t$ = -0.729, $p$-value = 0.466, 95\% CI: [-4.994; 2.288]). As with the other experiments, the results remain stable when we conduct subgroup analyses.

@@

Regulating violence is one of the most fundamental responsibilities of the
state, and in many places in the world today, states are failing to do so.
Brazil provides a telling example of how extrajudicial violence can escalate
and pose a serious threat to the stability of government institutions. From
2011 to 2015, the country registered about 2,500 lynching episodes, and 173
people were killed by angry mobs in 2015 alone---nearly one execution every two
days \citep{barbara2015vigilantes, oliveira2016mob}. According to José de Souza
Martins \citeyearpar{martins2015linchamentos}, who has studied lynchings in
Brazil for more than thirty years, these figures are not only the highest in
the country's history, but among the highest in the world. The people who
participate in lynchings are typically young men, but they also sometimes
include teenage women and girls, elderly women, and even members of the local
police \citep{moura2017linchamentos}. In this respect, interventions that
reduce popular support for extrajudicial violence can have a substantial impact
on citizen welfare and potentially increase state legitimacy in young
democracies.

Our results show that the characteristics of a lynching significantly effect
people's support for it, especially when the crimes are heinous or committed
against vulnerable populations. In contrast with studies about mob violence in
the American South, we find no evidence that lynchings are strictly racially
motivated in Brazil. While we were able to reduce support for lynchings by
reminding respondents that lynchings are a criminal act and that lynchings risk
igniting vendettas, appeals to human rights had no effect in our experiment.

>>>>

% English
\usepackage[english]{babel}
\usepackage[english]{isodate}
\cleanlookdateon

@@

% Bibliography
\usepackage{natbib}
\bibliographystyle{apalike}
\makeatletter
% Remove comma after author
\setcitestyle{aysep={}}
\patchcmd{\NAT@citex}
	  {\@citea\NAT@hyper@{%
		 \NAT@nmfmt{\NAT@nm}%
		 \hyper@natlinkbreak{\NAT@aysep\NAT@spacechar}{\@citeb\@extra@b@citeb}%
		 \NAT@date}}
	  {\@citea\NAT@nmfmt{\NAT@nm}%
	   \NAT@aysep\NAT@spacechar\NAT@hyper@{\NAT@date}}{}{}
	\patchcmd{\NAT@citex}
	  {\@citea\NAT@hyper@{%
		 \NAT@nmfmt{\NAT@nm}%
		 \hyper@natlinkbreak{\NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi}%
		   {\@citeb\@extra@b@citeb}%
		 \NAT@date}}
	  {\@citea\NAT@nmfmt{\NAT@nm}%
	   \NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi\NAT@hyper@{\NAT@date}}
	  {}{}
% Patch case where name and year are separated by aysep
\patchcmd{\NAT@citex}
  {\@citea\NAT@hyper@{%
     \NAT@nmfmt{\NAT@nm}%
     \hyper@natlinkbreak{\NAT@aysep\NAT@spacechar}{\@citeb\@extra@b@citeb}%
     \NAT@date}}
  {\@citea\NAT@nmfmt{\NAT@nm}%
   \NAT@aysep\NAT@spacechar\NAT@hyper@{\NAT@date}}{}{}
% Patch case where name and year are separated by opening bracket
\patchcmd{\NAT@citex}
  {\@citea\NAT@hyper@{%
     \NAT@nmfmt{\NAT@nm}%
     \hyper@natlinkbreak{\NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi}%
       {\@citeb\@extra@b@citeb}%
     \NAT@date}}
  {\@citea\NAT@nmfmt{\NAT@nm}%
   \NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi\NAT@hyper@{\NAT@date}}
  {}{}
\makeatother

@@

% Abstract
\begin{abstract}
\doublespacing \noindent Why do people support extrajudicial violence? In two survey experiments with respondents in Brazil, we examine which characteristics of lynching scenarios garner greater support for lynching and whether providing different types of information about lynching reduces support for it. We find that people often do support community members to take vengeance. In particular, our analysis finds that people strongly support the use of extrajudicial violence by families of victims against men who sexually assault and murder women and children. We also find that criminal punishment and the threat of vendettas reduce support, but appeals to the human rights of victims have zero effect on support for lynchings. Unlike the U.S. experience with lynchings, race was not observed to play an important role in how respondents answered the survey.
\vspace{.25cm}

\noindent \textbf{Keywords:} Brazil, crime, extralegal violence, lynching, vigilantism
\vspace{.25cm}

@@

The regulation of violence is a crucial function of the state. When states
control violence effectively, markets, politics, and civil society can flourish
\citep{besley2011pillars, north2009violence}. However, states still struggle to
maintain the monopoly of force in much of the Global South. Scholars have
written extensively about how rebel organizations build support to challenge
government institutions, yet we know little about what motivates local
extrajudicial violence and vigilantism in the developing world
\citep{bateson2020politics}.\footnote{On vigilantism more generally, see
\citet{schuberth2013challenging,
zizumbo2017community,smith2019contradictions,cohen2022collective}.} This
omission is significant not only because violence is pervasive in weak states,
but it also comes at a great cost \citep{blair2017predicting}.

One of the most serious forms of extrajudicial violence is lynching, which can
be defined as ``incidents of physical violence committed by large numbers of
private citizens against one or more individuals accused of having committed a
``criminal'' offense, whether or not this violence resulted in the death of the
victim(s)'' \citep[645]{godoy2004justice}. Although lynchings occur in more
than one hundred countries in all regions of the world
\citep{jung2020lynching,smith2019contradictions}, Latin America has been
particularly affected by a sharp increase in vigilante violence. Lynching
episodes have been reported in Guatemala, Mexico, Peru, Venezuela and other
countries in the region \citep{cruz2019determinants, godoy2004justice,
barbara2015vigilantes}. From 2011 to 2015, Brazil registered about 2,500
lynching episodes, and 173 people were killed by angry mobs in 2015
alone---nearly one execution every two days \citep{barbara2015vigilantes,
oliveira2016mob}. According to José de Souza Martins
\citeyearpar{martins2015linchamentos}, who has studied lynchings in Brazil for
more than thirty years, these figures are not only the highest in the country's
history, but among the highest in the world. The people who participate in
lynchings are typically young men, but they also sometimes include teenage
women and girls, elderly women, and even members of the local police
\citep{moura2017linchamentos}.

These alarming figures suggest that, despite being dismissed by policymakers as
irrational and senseless, vigilantism enjoys considerable approval in Latin
America \citep{berg2011globalizing, goldstein2005flexible}. This raises a
number of puzzles: under which conditions do citizens perceive lynchings as a
legitimate response to crime? Are some groups seen as more ``deserving'' of
lynchings that others? And how can we reduce pro-vigilantism attitudes in
young, unequal democracies?

In this study, we run two survey experiments and use primary qualitative data
to understand attitudes toward lynchings in Brazil. Our paper begins with a
conjoint experiment that evaluates in which circumstances citizens perceive
lynchings as a justifiable punishment for crime. We also included open-ended
questions for respondents to elaborate on their decision process. Lastly, we
employ an information provision experiment to test the efficacy of three
strategies to reduce individual support for lynchings.

Our empirical approach offers several distinct advantages. First, the data we
collected come from the same level of analysis as the phenomenon itself: local
residents. Second, online surveys are less susceptible to social desirability
bias, and survey experiments help to elicit sensitive information---an
especially important benefit in researching this topic
\citep{horiuchi2022does}. Third, the experimental nature of the approach allows
us to precisely estimate the size of effects and compare different theoretical
explanations simultaneously. Finally, we can also control for confounding,
which is a potential concern in previous research on vigilantism.

@@

In addition to our knowledge of the Brazilian case, we also chose our
attributes based partly on existing work on crime, vigilantism, and
extrajudicial violence. From the prison violence literature, we know that the
relative age differences and the genders of perpetrators and victims affect how
it is perceived \citep{fleisher2009myth}. We likewise know that certain
offenses are seen as especially reprehensible \citep{skarbek2014social}, so we
included such offenses, like molestation, alongside less serious offenses. We
include race because, in the United States, violent lynchings have often been
wielded for racist reasons \citep{dray2003hands}. Likewise, we include
residency because extrajudicial violence is often used against people ``who
don't belong'' or are ``outsiders'' in some respect. Finally, in honor cultures
that value retaliatory violence, people believe that victims and the family of
victims have a special right (and often an obligation) to enact retribution
\citep{weiner2013rule}.

@@

We report marginal means in our main analysis instead of average marginal
component effects (AMCEs). \citet{leeper2020measuring} show that AMCEs can be
misleading in subgroup comparisons as model results are sensitive to the choice
of reference categories in interactions. In contrast, marginal means allow for
a clear description of quantities of interest, in our case preferences about
lynching, while also allowing for easy comparisons between groups of
respondents. In a forced-choice experiment such as this one, a marginal-means
estimate of 0.5 indicates that respondents are indifferent to this attribute
vis-à-vis other attributes. When the coefficient is lower than 0.5, respondents
dislike profiles with this attribute. Conversely, when the point estimate is
higher than 0.5, respondents prefer profiles containing a given attribute.

@@

Respondents selected male offenders as preferred lynching victims more often
than female offenders. We found identical results in our subgroup analyses, in
which we disaggregated our sample according to respondents' gender, income,
race, support for the death penalty, and views on the police and judiciary. In
all cases, men and teenage boys were more likely to be selected as more
justified victims of lynching. Compared with the attributes we discuss below,
the gender of the crime perpetrator had one of the largest effects in our
experiment. With regard to the age of the crime perpetrator, respondents
tolerated the lynching of adults more than attacks against teenagers or the
elderly. More precisely, subjects were indifferent toward teenagers but reacted
positively to lynching of adults and negatively to lynching of elderly people.

When we analyzed the effect of the race of the crime perpetrator, in contrast
with the American experience, respondents favored the lynching of white
perpetrators the most. Of all the racial identities that we included in the
experiment, Black people were the least likely to be chosen as a more justified
lynching victim, although the 95 percent confidence intervals overlap with the
remaining categories. The effect is also robust for most subgroups, including
white respondents. Asian respondents, however, rated fellow Asians as their
least preferred choice. Taken together, the results provide experimental
evidence that support for lynchings in Brazil does not resemble the typical
racial patterns scholars have observed in the United States
\citep{dray2003hands,seguin2019national,obert2018keeping}. These findings are
consistent with recent research on vigilantism in Haiti
\citep{jung2020lynching} and with journalistic observations. They further
suggest that the US experience with lynchings might be distinct from
vigilantism in other places and times \citep{oliveira2016mob}.

The residency of the perpetrator is not a significant driver of support for
lynching. Subjects were indifferent as to whether the crime perpetrator lives
in the area or in another neighborhood. In this sense, it is possible that
subjects see lynchings not as a means to protect the area in which they live,
which people commonly suppose vigilante groups do, but rather as retribution in
specific cases.

The next attribute is type of offense. Rape was by far the most significant
attribute in our model, followed by murder and molestation. It is not
surprising that murder and rape were the most important offenses to
respondents. Overall, we found that respondents do not believe that property
crimes, such as car theft and pickpocketing, provide justification for
lynching. When we analyzed the data by groups, we also saw that poorer
respondents, whose household income is less than R\$3,000 (US\$650) per month,
are more likely to support the lynching of individuals who commit a property
crime.

For age, we observed that respondents were particularly concerned about
children as crime victims, as this attribute is strongly correlated with
support for lynchings. Respondents also thought lynchings are more justified
when the victim is a teenager, an elderly person, or an adult.

When we analyzed the gender of the crime victim, we found that respondents
support lynchings more strongly when the victim is a woman or teenage girl and
less strongly when the lynching victim is a man or teenage boy. The difference
here is smaller than when we considered the gender of the crime perpetrator,
but as in the previous estimation the 95 percent confidence intervals of each
attribute did not overlap.

Lastly, and related to this, we examined who respondents believed are more
legitimate perpetrators of lynchings. Respondents considered the family to be
the most justified in carrying out a lynching. Again, the intervals here did
not overlap with those of any other category. Bystanders, gangs, and neighbors
do not increase or decrease respondents attitudes about lynching. If the police
carry out a lynching, that reduced the perceived justification. In sum,
respondents did not believe that lynchings should be carried out by the state
but did believe they should be used as a tool for individual or family
retribution. These results are consistent with norms of an honor culture in
which offenses are seen to tarnish the victim's status and the only way to
remove the stigma is through self-help efforts in the form of retaliatory
violence \citep{nisbett2018culture}.

@@

\begin{table}[ht] \centering
  \caption{Average treatment effects for experiment 2}
  \label{tab:exp02}
\begin{tabular}{@{\extracolsep{3pt}}lD{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3} }
\\[-1.8ex]\hline \\[-1.8ex]
\\[-1.8ex] & \multicolumn{4}{c}{\textbf{Lynching Support}\vspace{.5cm}} \\
\\[-1.8ex] & \multicolumn{1}{c}{(1)} & \multicolumn{1}{c}{(2)} & \multicolumn{1}{c}{(3)} & \multicolumn{1}{c}{(4)}\\
\hline \\[-1.8ex]
 Constitution and penal code & -4.509^{**} &  &  &  \\
  & (1.805) &  &  &  \\
  Human rights &  & -1.571 &  &  \\
  &  & (1.801) &  &  \\
  Vendettas &  &  & -3.156^{*} &  \\
  &  &  & (1.879) &  \\
  Combined treatments &  &  &  & -3.023^{**} \\
  &  &  &  & (1.493) \\
  Constant & 40.823^{***} & 40.823^{***} & 40.823^{***} & 40.823^{***} \\
  & (1.293) & (1.293) & (1.293) & (1.293) \\
 N & \multicolumn{1}{c}{1,114} & \multicolumn{1}{c}{1,173} & \multicolumn{1}{c}{1,092} & \multicolumn{1}{c}{2,215} \\
\hline \\[-1.8ex]
\multicolumn{5}{l}{$^{*}$p $<$ .1; $^{**}$p $<$ .05; $^{***}$p $<$ .01} \\
\multicolumn{5}{l}{Robust standard errors in parentheses.} \\
\end{tabular}
\end{table}
\normalsize

Table \ref{tab:exp02} summarizes our main results for this experiment. We find
that when respondents are reminded of legal punishments associated with
lynchings, their support for lynching decreases by about 11 percent. This
result is consistent with the idea that respondents will support an activity
less if they perceive it to be more costly. The next treatment group was told
the human rights message. The result is not statistically significant and
indicates that appeals to individual liberties and the rule of law are likely
to fall short. This finding has practical implications, as many campaigns
against violent crimes are conducted by human rights groups, which tend to
emphasize this type of message. However, participants do not generally believe
that offenders do, in fact, have human rights. Accordingly, we have little
reason to believe that such messaging is effective. By contrast, Treatment 1
reminds respondents that perpetrators of lynching potentially face a high
personal cost from doing so.

Our last treatment, which raises the possibility of sparking vendettas, has a
large, negative effect on support for lynchings. Mentioning that lynchings
trigger cycles of violence has a negative effect of about 8 percent when
compared with the control group. While we expected that the coefficient for
this treatment condition would be larger than that for the treatment on legal
punishments, the means of both groups are not statistically different from each
other. As with the other experiments, the results remain stable when we conduct
subgroup analyses.

@@

Regulating violence is one of the most fundamental responsibilities of the
state, and in many places in the world today, states are failing to do so.
Brazil provides a telling example of how extrajudicial violence can escalate
and pose a serious threat to the stability of government institutions.

%%%%DELETE OR REVISE In this respect, interventions that reduce popular support for extrajudicial violence can have a substantial impact on citizen welfare and potentially increase state legitimacy in young democracies.

Our results show that the characteristics of a lynching significantly affect
people's support for it, especially when the crimes are heinous or committed
against vulnerable populations. In contrast with studies about mob violence in
the American South, we find no evidence that lynchings are strictly racially
motivated in Brazil. While we were able to reduce support for lynchings by
reminding respondents that lynchings are a criminal act and that lynchings risk
igniting vendettas, appeals to human rights had no effect in our experiment.
